---
title: "R para Data Science"
output: html_document
date: '2022-06-09'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls(all=TRUE))
```

1.  Visualização de dados com o ggplot2

```{r}
## instalando pacote
if(c("ggplot2") %in% rownames(installed.packages())==FALSE)
{
  install.packages(c("ggplot2"), dependencies=TRUE)
}
library("ggplot2")

if("dados" %in% rownames(installed.packages())==FALSE)
{
  install.packages("dados")
}
library("dados")
```

Criando um ggplot

```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy))
```

Exercícios

1.  Run ggplot(data = mpg). What do you see?

```{r}
ggplot(data = mpg)
```

R: aparece um gráfico em branco, pois o ggplot2 funciona a partir de camadas, neste caso teria que acrescentar mais informações como o tipo de gráfico "geom".

2.  How many rows are in mtcars? How many columns?

```{r}
dim(mtcarros)
```

R: o data frame mtcars possui 32 linhas/observações e 11 colunas/variáveis

3.  What does the drv variable describe? Read the help for ?mpg to find out.

```{r}
?mpg
```

R: tipo de tração

4.  Make a scatterplot of hwy versus cyl.

```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x = hwy, y = cyl))
```

5.  What happens if you make a scatterplot of class versus drv? Why is the plot not useful?

```{r}
ggplot(data=mpg)+
  geom_point(mapping = aes(x = class, y = drv))
```

R: o gráfico não consegue, de forma clara, expressar as informações que gostaríamos de vê.

Mapeamoentos Estéticos

```{r}
# classe de cada carro = cores dos pontos
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy, colour = class))
```

```{r}
# estética dos pontos = tamanho dos pontos 
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```

```{r}
# estética transparencia dos pontos
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```

```{r}
# shape dos pontos
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

```{r}
# todos os pontos azuis
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

Exercises

1.  What's gone wrong with this code? Why are the points not blue?

```{r}
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

R: porque a cor "blue" não foi configurada como um argumento da função geom, isto é, fora do aes().

2.  Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset.) How can you see this information when you run mpg?

```{r}
# R
str(mpg)
??mpg
```

R: variáveis categóricas (manufacturer, model, trans, drv, fl, class); variáveis continuas (displ).

3.  Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical versus continuous variables?

```{r}
# color
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy))+
  aes(color = displ)
```

```{r}
# size
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy))+
  aes(size = displ)
```


4.  What happens if you map the same variable to multiple aesthetics?

```{r}
# shape não pode ser usado para variáveis contínuas
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy),color = "blue", size = 10)
```

R: ele vai aplicar uma combinação de todas as estéticas ao gráfico.

5.  What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point.)

```{r}
?geom_point
ggplot(data = mpg)+
  geom_point(mapping = aes(x = displ, y = hwy)) + aes(stroke = displ)
```
6. What happens if you map an aesthetic to something other than a variable name, like aes(color = displ < 5)?
```{r}
ggplot(data = mpg)+
  geom_point(aes(x = displ, y = hwy))+
  aes(color = displ < 5)
```
R: a cor só foi alterada para os valores de displ menores do que 5.

Criando facetas de um gráfico a partir de uma única variável

```{r}
ggplot(data = mpg)+
  geom_point(aes(x = displ, y = hwy))+
  facet_wrap(~class, nrow = 2)
```

Criando faceta de um gráfico a partir de duas variáveis.

```{r}
ggplot(data = mpg)+
  geom_point(aes(x = displ, y = hwy))+
  facet_grid(~drv ~cyl)
```

